Composable Generalized Consensus

### 2015-07-21 paper/intro.tex
A few fixes.

Allowing the execution of commuting commands to proceed without
synchronization is an important practical optimization, as commands may
often commute in practice, for example when operating on disjoint parts
of a data structure. To enable the synchronization-free execution of
commuting commands, we have based our framework on Generalized Consensus
\cite{Lamport05GeneralizeConsensus}. Traditional SMR implementations use a
separate instance of a consensus algorithm to agree on the command to execute
at each index of the command sequence on which the replicas must agree. Using
separate consensus instances makes it hard to optimize the execution of
commuting command because commutativity is not a property of a single command.
Instead of restricting agreement to a single value, the Generalized Consensus
problem requires that set of replica servers agree on the state of a growing
data structure called a command-structure set. A command-structure set is
a data structure with an append operation such that the order in which two
commuting commands are appended does not change the resulting state. Thanks to
this property, two replicas executing commuting commands in a different order
will not end up in a different state and can continue executing without further
synchronization. In message-passing systems, this property can be leveraged to
reduce the latency of processing one command to two message delays instead of
three. 
%As we will see, fast execution of commutative requests also enables the processing of read-only commands
%in two message delays, although read-only commands do not commute with write
%commands.

In practice, a SMR algorithm must support fast reconfiguration to replace
faulty components, without which the long-term stability of a system is
inevitably compromised. Our framework abstracts over how the reconfiguration
decisions are made, providing an interface allowing to plug in an external
reconfiguration master implementing the reconfiguration policy of the
system. The master decides which GC module to run next and which replica
servers will participate, removing failed replicas and adding new ones. The
reconfiguration master enables tolerating $k$ failures with $k+1$ replicas
\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}, an important
improvement over the traditional $2k+1$ needed replicas.

>>>>

\begin{comment}
Allowing the execution of commuting commands to proceed without
synchronization is an important practical optimization, as commands may
often commute in practice, for example when operating on disjoint parts
of a data structure. To enable the synchronization-free execution of
commuting commands, we have based our framework on Generalized Consensus
\cite{Lamport05GeneralizeConsensus}. Traditional SMR implementations use a
separate instance of a consensus algorithm to agree on the command to execute
at each index of the command sequence on which the replicas must agree. Using
separate consensus instances makes it hard to optimize the execution of
commuting command because commutativity is not a property of a single command.
Instead of restricting agreement to a single value, the Generalized Consensus
problem requires that set of replica servers agree on the state of a growing
data structure called a command-structure set. A command-structure set is
a data structure with an append operation such that the order in which two
commuting commands are appended does not change the resulting state. Thanks to
this property, two replicas executing commuting commands in a different order
will not end up in a different state and can continue executing without further
synchronization. In message-passing systems, this property can be leveraged to
reduce the latency of processing one command to two message delays instead of
three. 
%As we will see, fast execution of commutative requests also enables the processing of read-only commands
%in two message delays, although read-only commands do not commute with write
%commands.

In practice, a SMR algorithm must support fast reconfiguration to replace
faulty components, without which the long-term stability of a system is
inevitably compromised. Our framework abstracts over how the reconfiguration
decisions are made, providing an interface allowing to plug in an external
reconfiguration master implementing the reconfiguration policy of the
system. The master decides which GC module to run next and which replica
servers will participate, removing failed replicas and adding new ones. The
reconfiguration master enables tolerating $k$ failures with $k+1$ replicas
\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}, an important
improvement over the traditional $2k+1$ needed replicas.
\end{comment}

### 2015-07-21 paper/paper.tex
A few fixes.

\vspace{2cm}
\begin{center}
This is a regular paper.
\end{center}

@@

By definition, all the sequences represented by a given history contain the same commands
and determine the same outputs for each command. Therefore, to implement a
replicated service, it is sufficient that learners learn histories instead of
sequences. For a thorough discussion of command histories we refer the reader
to \cite{Lamport05GeneralizeConsensus}. 
Our Isabelle/HOL formalization abstracts over the concrete representation of histories, assuming only their
necessary properties. Command histories were first introduced in trace theory
\cite{Mazurkiewicz84Semantics}.

@@

@@

\subsection{The Leader Algorithm}

@@

Instead, we will now show how to obtain an algorithm with performance characteristics close to those of Fast Paxos, employing a leader to deal with contention, without modifying at all our existing Broadcast-Decide algorithm.
To do so, we define a new kind of rounds, the Leader rounds, which uses the same inter-round interface as the Broadcast-Decide algorithm. 
The inter-round interface will allow use to combine Leader rounds and Broadcast-Decide rounds in any order, obtaining an algorithm which can adapt its strategy to its environment, selecting the Broadcast-Decide algorithm under low contention and switching to the slower Leader algorithm when there is contention.

@@

Composable Generalized Consensus introduces two new transitions: $Init\left( s, h \right)$ and $Abort\left( s, h
\right)$, for a switcher $s$ and a history $h$. The transition $Init\left( s,h \right)$ inserts $h$ in the set $inits$, and models the switcher $s$ receiving
the abort history $h$ from the preceding round. The transition $Abort\left( s, h \right)$ inserts $h$ in the set $aborts$ and models the switcher $s$ passing the abort history $h$ to the next round.  
All the transition are guarded by the following conditions:
\begin{compactitem}
    \item[\textbf{Agreement}:] The set $\aset{\bigcup learned'\left[ l
      \right] : l \in Learners}$ is compatible.
    \item[\textbf{Validity}:] There is a history $h$ belonging to $inits$ such that every abort history or learned history 
        is of the form $h \star cs$, where $cs$ contains only commands
        of the set $proposed$.
    \item[\textbf{Initialization}:] If $inits$ is empty, then no
        history can be learned and no switcher can abort.
    \item[\textbf{Irrevocability}:] In the case of a $Learn\left( l,h
        \right)$ transition, the new learned history $h$ is an 
        extension of every history in $learned\left[ l \right]$.
    \item[\textbf{Safe Abort}:] For every abort history $h$ there is a learner $l$ such that $h$ is an extension of the maximum learned history of $l$.
\end{compactitem}

@@

We provide a mechanically-checked proof of the composition theorem in Isabelle/HOL at \url{http://losa.fr/cgc}.
The proofs uses a version of the specification of CGC that has a more
executable form that presented in this section. 
The Isabelle/HOL formalization of CGC uses the theory of I/O automata, which
is also formalized in the Isabelle/HOL. The proof uses a refinement
mapping from the composition of two consecutive CGC instances to a single CGC
instance.

@@

The rounds of Classic Paxos are similar to the Leader algorithm of section 4.2 but
tolerates that a minority of servers crash by using majority quorums instead of
waiting for histories from all the acceptors.

>>>>

%\vspace{2cm}
%\begin{center}
%This is a regular paper.
%\end{center}

@@

By definition, all the sequences represented by a given history contain the same commands
and determine the same outputs for each command. Therefore, to implement a
replicated service, it is sufficient that learners learn histories instead of
sequences.

@@

For a thorough discussion of command histories we refer the reader
to \cite{Lamport05GeneralizeConsensus}, where histories are called C-structs. 
%Our Isabelle/HOL formalization abstracts over the concrete representation of histories, assuming only their
%necessary properties. 
Command histories were first introduced in trace theory by Mazurkiewicz 
\cite{Mazurkiewicz84Semantics}.

@@

\subsection{The Leader Algorithm} 
\label{sec:leader}

@@

Instead, we will now show how to obtain an algorithm with performance characteristics close to those of Fast Paxos, employing a leader to deal with contention, without modifying at all our existing Broadcast-Decide algorithm.
To do so, we define a new kind of rounds, the Leader rounds, which uses the same inter-round interface as the Broadcast-Decide algorithm. 
The inter-round interface will allow us to combine Leader rounds and Broadcast-Decide rounds in any order, obtaining an algorithm which can adapt its strategy to its environment, selecting the Broadcast-Decide algorithm under low contention and switching to the slower Leader algorithm when there is contention.

@@

Composable Generalized Consensus introduces two new transitions: $Init\left( s, h \right)$ and $Abort\left( s, h
\right)$, for a switcher $s$ and a history $h$. The transition $Init\left( s,h \right)$ inserts $h$ in the set $inits$, and models the switcher $s$ receiving
the abort history $h$ from the preceding round. The transition $Abort\left( s, h \right)$ inserts $h$ in the set $aborts$ and models the switcher $s$ passing the abort history $h$ to the next round.  
All the transition are guarded by the following conditions:
\begin{compactitem}
    \item[\textbf{Agreement}:] The set $\aset{\bigcup learned'\left[ l
      \right] : l \in Learners}$ is compatible.
    \item[\textbf{Validity}:] There is a history $h$ belonging to $inits$ such that every abort history or learned history 
        is of the form $h \star cs$, where $cs$ contains only commands
        of the set $proposed$.
    \item[\textbf{Initialization}:] If $inits$ is empty, then no
        history can be learned and no switcher can abort.
    \item[\textbf{Irrevocability}:] In the case of a $Learn\left( l,h
        \right)$ transition, the new learned history $h$ is an 
        extension of every history in $learned\left[ l \right]$.
    \item[\textbf{Safe Abort}:] Every abort history $h$ is an upper bound of the set of learned histories.
\end{compactitem}

@@

We provide a mechanically-checked proof of the composition theorem in Isabelle/HOL at \url{http://losa.fr/cgc}.
The proofs uses a version of the specification of CGC that has a more
executable form that presented in this section. 
The Isabelle/HOL formalization of CGC uses the theory of I/O automata, which
is also formalized in Isabelle/HOL. The proof uses a refinement
mapping from the composition of two consecutive CGC instances to a single CGC
instance.

@@

The rounds of Classic Paxos are similar to the Leader algorithm of \cref{sec:leader} but
tolerate that a minority of servers crash by using majority quorums instead of
waiting for histories from all the acceptors.
